---
title: "SLR: Indicator Variables"
author: "Your Name"
date: "date here"
output: html_document
---


```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(openintro)
```


## Indicator Variables

In regression analysis, explanatory variables are treated as a numerical. What if we want to include a categorical variable?

**______________________**, also known as dummy variables, take on the values of `0` or `1` to represent the presence or absence of an attribute of a categorical variable. 

- Dummy variables allow us to convert categorical variables into numerical ones. 
- For example, a logical variable that takes the values, TRUE or FALSE are assigned the values 1 for TRUE and 0 for FALSE. 

For more information and examples of categorical variables with more than 2 categories, see https://www.sjsu.edu/faculty/guangliang.chen/Math261a/Ch8slides-indicator-variables.pdf

.footnote[Source: https://en.wikipedia.org/wiki/Dummy_variable_(statistics)]



## Simple Linear Regression model

$y_i = \beta_0 +\beta_1x_i + \epsilon_i$

$\beta_0$ is y-intercept  
$\beta_1$ is slope  
$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point

The explanatory variable, $x_i$, can either be **_____________** or **__________________**.

In this case, we will let $x_i$ be **categorical with 2 categories.** 

---

#### Data `babies` in `openintro` package

```{r echo = FALSE}
glimpse(babies)
```



This time we are going to look at the linear relationship between `birth weight`, as the response, and `smoke`, as the explanatory variable. 


| y | Response    | Birth weight | Numeric |
|---|-------------|-----------------|---------|
| x | Explanatory | Smoke           | Categorical |




The variable `smoke` indicates whether the mother of a baby smoke during her pregnancy. Note that the variable is assigned as `integer`, and hence we have 1 for yes and 0 for no. 

Even is this variable was converted to `logical`, the model would still treat the variable as "integer", assigning 1 for TRUE and 0 for FALSE when calculating the slope and y-intercept. 

```{r}
babies |> 
select(bwt, smoke)
```

### Scatter plot of smoke vs birthweight
```{r}

```


### Finding the least-squares regression line

Recall, that to find the coefficients $b_0$ and $b_1$ of the model $\hat {y}_i = b_0 + b_1 x_i$, we use the function `lm(y ~ x, data = data_name)`.

```{r}

```



$\hat {y}_i = b_0 + b_1 x_i$


$\hat {\text{bwt}_i} = b_0 + b_1 \text{ smoke}_i$

$\hat {\text{bwt}_i} = _______ + (________\text{smoke}_i)$


---

#### Expected bwt for a baby with a non-smoker mother

x = ________

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$

$\hat {\text{bwt}_i} = 123 + (-8.94\times 0)$

$\hat {\text{bwt}_i} = _________$

$E[bwt_i | smoke_i = 0] = b_0$

Interpretation: _______


#### Expected bwt for a baby with a smoker mother

x = _________

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$

$\hat {\text{bwt}_i} = 123 + (-8.94\times 1)$

$\hat {\text{bwt}_i} = ___________$

$E[bwt_i | smoke_i = 1] = b_0 + b_1$

Interpretation: _____________


Since we are using a categorical variable, `smoke`, individuals are then classified under one of two categories: non-smoker mother vs smoker mother. That means we have two groups. 


- Difference between groups: _______________________

When the predictor variable is categorical, with two categories the **slope**, $b_1$, is no longer described as a rate of change, but as the **expected difference between the two groups.**

---

#### Plotting the regression line on the scatter plot

```{r}

```


---

## Inference: Confidence Intervals

Recall, that we can use sample data to construct a confidence interval to estimate the parameters, $\beta_1$ or $\beta_0$. 

Use the function `confint(model, level = CL)`, with the linear model as the first argument, followed by `level = 0.90, 0.95, or 0.99` to indicate the level of confidence we want to use. 


```{r}

```



### Interval for slope $\beta_1$

Interval: (_____ , ______)

Note that the confidence interval for the "slope" does not contain 0 and all the values in the interval are negative. 

We can't say that smoke is the cause for a lower birth weight, as there may be other factors that may cause a mother to smoke, such as stress. However, we can say that smoke has a negative relationship with birth weight.

We are 95% confident that ____________________.




### Interval for y-intercept $\beta_0$

Interval: (______ , ______)

Note that, in this example, the y-intercept is interpretable. That is, $\hat{y} = 123$ when $x = 0$ indicates the birth weight for a baby whose mother did not smoke during pregnancy. Hence an interval would also make sense. 

**We are 95% confident to conclude that ________________________.


