---
title: "SLR: Indicator Variables"
author: "Alma Castro, Cypress College"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
    anchor_sections: FALSE
---

class: title-slide


<br>
<br>
.pull-right[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(openintro)
library(broom)

```


## Indicator Variables

In linear regression analysis, explanatory variables are treated as numerical. What if we want to include a categorical variable?

**Indicator variables**, also known as dummy variables, take on the values of `0` or `1` to represent the presence or absence of an attribute of a categorical variable. 

- Dummy variables allow us to convert categorical variables into numerical ones. 
- For example, a logical variable that takes the values, TRUE or FALSE are assigned the values 1 for TRUE and 0 for FALSE. 

For more information and examples of categorical variables with more than 2 categories, see https://www.sjsu.edu/faculty/guangliang.chen/Math261a/Ch8slides-indicator-variables.pdf

.footnote[Source: https://en.wikipedia.org/wiki/Dummy_variable_(statistics)]

---

## Simple Linear Regression model

$y_i = \beta_0 +\beta_1x_i + \epsilon_i$

$\beta_0$ is y-intercept  
$\beta_1$ is slope  
$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point

The explanatory variable, $x_i$, can either be **numerical** or **categorical**.

In this case, we will let $x_i$ be categorical with 2 categories. 

---

#### Data `babies` in `openintro` package

```{r echo = FALSE}
glimpse(babies)
```

---

This time we are going to look at the linear relationship between `birth weight`, as the response, and `smoke`, as the explanatory variable. 


| y | Response    | Birth weight | Numeric |
|---|-------------|-----------------|---------|
| x | Explanatory | Smoke           | Categorical |


.pull-left[
The variable `smoke` indicates whether the mother of a baby smoke during her pregnancy. Note that the variable is assigned as `integer`, and hence we have 1 for yes and 0 for no. 

Even is this variable was converted to `logical`, the model would still treat the variable as "integer", assigning 1 for TRUE and 0 for FALSE when calculating the slope and y-intercept.
]

.pull-right[
```{r}
babies |> 
select(bwt, smoke)
```
]

---

### Scatter plot of smoke vs birthweight

.pull-left[
```{r eval = FALSE}
babies |> 
  ggplot(
    aes(x = smoke, 
        y = bwt)) +
  geom_point()
```
]

.pull-right[
```{r echo = FALSE, warning = FALSE}
babies |> 
  ggplot(aes(x = smoke, y = bwt)) +
  geom_point()
```

]


---

### Finding the least-squares regression line

Recall, that to find the coefficients $b_0$ and $b_1$ of the model $\hat {y}_i = b_0 + b_1 x_i$, we use the function `lm(y ~ x, data = data_name)`.

```{r}
model_s <- lm(bwt ~ smoke, data = babies)
model_s
```


---

### Least-squares Regression Line

$\hat {y}_i = b_0 + b_1 x_i$

$\hat {\text{bwt}_i} = b_0 + b_1 \text{ smoke}_i$

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$


---

.pull-left[

#### Expected bwt for a baby with a non-smoker mother

**x = smoke = 0**

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$

$\hat {\text{bwt}_i} = 123 + (-8.94\times 0)$

$\hat {\text{bwt}_i} = 123$


The expected birth weight of a baby with a non-smoker mother is 123 ounces. 

]

--

.pull-right[
#### Expected bwt for a baby with a smoker mother

 **x = smoke = 1**.

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$

$\hat {\text{bwt}_i} = 123 + (-8.94\times 1)$

$\hat {\text{bwt}_i} = 114.06$


The expected birth weight of a baby with a smoker mother is 114 ounces.

]

---

.pull-left[

#### Expected bwt for a baby with a non-smoker mother


$E[bwt_i | smoke_i = 0] = 123 = b_0$


#### Expected bwt for a baby with a smoker mother

$E[bw t_i | smoke_i = 1] = 114$ 
$= 123 + (-8.94)$ 
$= b_0 + b_1$
]

--

.pull-right[

Since we are using a categorical variable, `smoke`, babies are then classified under one of two categories: non-smoker mother vs smoker mother. That means we have two groups.

- Difference between groups: 123 - 114 = -9 which is the $b_1$ (slope)

When the predictor variable is categorical, with two categories the **slope**, $b_1$, is no longer described as a rate of change, but as **the expected difference between the two groups.**

]

---

### Plotting the regression line on the scatter plot

.pull-left[
```{r eval = FALSE, message = FALSE, warning = FALSE}
ggplot(babies,
         aes(x = smoke, 
             y = bwt)) +
  geom_point() +
  geom_smooth(method = "lm", 
              se = FALSE)
```
]

.pull-right[

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(babies,
         aes(x = smoke, y = bwt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

]

---
## Inference: Confidence Intervals

Recall, that we can use sample data to construct a confidence interval to estimate the parameters, $\beta_1$ or $\beta_0$. 

Use the function `confint(model, level = CL)`, with the linear model as the first argument, followed by `level = 0.90, 0.95, or 0.99` to indicate the level of confidence we want to use. 


```{r}
confint(model_s, level = 0.95)
```

---

### Interval for slope $\beta_1$

(-10.96  -6.91)

Note that the confidence interval for the "slope" does not contain 0 and all the values in the interval are negative. 

We can't say that smoke is the cause for a lower birth weight, as there may be other factors that may cause a mother to smoke, such as stress. However, we can say that smoke has a negative relationship with birth weight.

--

In fact, we are 95% confident that the expected difference in birth weight for babies of mothers who smoke and don't smoke is between 7 to 11 ounces. 

More specifically, **the expected birth weight for babies of mothers who smoke is between 7 to 11 ounces less than babies of mothers who don't smoke.** 

---

### Interval for y-intercept $\beta_0$

(121.77 124.32)

Note that, in this example, the y-intercept is interpretable. That is, $\hat{y} = 123$ when $x = 0$ indicates the birth weight for a baby whose mother did not smoke during pregnancy. Hence an interval would also make sense. 

--

**We are 95% confident to conclude that the expected birth weight for babies of mothers who do not smoke during their pregnancy is between 121.8 and 124.3 ounces.**





