---
title: "SLR: Transformations"
author: "Alma Castro, Cypress College"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
    anchor_sections: FALSE
---

class: title-slide


<br>
<br>
.pull-right[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`

]

---


```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(broom)
library(openintro)

```

We know that in order to use a linear model there are certain conditions that needed to be met: 

  - linearity
  - constant variance
  - normality of residuals 

What if the condition of linearity is not met? 

--

We can transform the data so that the relationship gets close to a linear trend. 

---

#### Data `cars93` in `openintro` package

```{r}
cars <- cars93
glimpse(cars)
```

---
## Example

.pull-left[
Let's find a model to predict city miles per gallon using the weight of the car.

| y | Response    | mpg_city  | Numeric |
|---|-------------|---------|---------|
| x | Explanatory | weight | Numeric |


```{r fig.align='center', message=FALSE, eval = FALSE}
ggplot(cars, 
       aes(x = weight, 
           y = mpg_city)) +
  geom_point() +
  geom_smooth(se = FALSE)
  
```


The heavier the car, the fewer miles per gallon you get, but they do not have a linear association. It's curved!


]

.pull-right[
```{r fig.align='center', echo=FALSE, message=FALSE}
ggplot(cars, 
       aes(x = weight, 
           y = mpg_city)) +
  geom_point() +
  geom_smooth(se = FALSE)
  
```
]

---

### Using residual plot to verify linearity condition

- Run model:

```{r eval = FALSE, message = FALSE, warning = FALSE}
model_city <- 
  lm(mpg_city ~ weight, data = cars)
```

- Add residuals to data: 

```{r eval = FALSE, message = FALSE, warning = FALSE}
cars <- cars |> 
  modelr::add_residuals(model_city)
```

---

- Create residual plot:

.pull-left[
```{r eval = FALSE, message = FALSE, warning = FALSE}
ggplot(cars, 
       aes(x = weight, 
           y = resid)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_hline(yintercept = 0)
```


The residual plot reveals a clear curve trend, hence a **linear model is not appropriate**. 

What can we do then?
]

.pull-right[

```{r echo = FALSE, message = FALSE, warning = FALSE}
model_city <- 
  lm(mpg_city ~ weight, data = cars)
cars <- cars |> 
  modelr::add_residuals(model_city)
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(cars, 
       aes(x = weight, y = resid)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_hline(yintercept = 0)
```

]
---

## Transformations of response variable (*y*)

When the relationship between an explanatory variable *x* and the response variable *y* is not linear and instead follows a curve, then one thing we can do is transform the data of *y*. 

The transformed data then lessens the curve and approaches it closer to a linear relationship with the explanatory variable. 

---

### Log transformation of *y*

One of the most common transformations is the log transformation where we take the log of the response variable values, `log(y)`.

  - Note that log is natural log in R.
  - Data values cannot be negative. 
  - Check data for 0's and negative values. 
  - If there are a few 0's then we can add a small amount before taking the log, e. g. `log(y + 1)`.
  - If y has many zero values, or any negative values, a log-transformation might not be the best choice.



---

In our example, *y* = mpg_city, which is all positive, so we would take the log of `mpg_city` and plot it on the y-axis.

.pull-left[

```{r eval = FALSE}
ggplot(cars, 
       aes(x = weight, 
           y = log(mpg_city))) + #<<
  geom_point()
  
```
]

.pull-right[
```{r echo=FALSE}
ggplot(cars, 
       aes(x = weight, y = log(mpg_city))) +
  geom_point() +
  scale_x_continuous(labels = scales::comma_format()) 
  
```

]

---
.pull-left[

```{r echo = FALSE,message = FALSE}
ggplot(cars, 
       aes(x = weight, 
           y = mpg_city)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  scale_x_continuous(labels = scales::comma_format()) 
  
```
]

.pull-right[
```{r echo=FALSE, message = FALSE}
ggplot(cars, 
       aes(x = weight, y = log(mpg_city))) +
  geom_point() +
  geom_smooth(se = FALSE) +
  scale_x_continuous(labels = scales::comma_format()) 
  
```

]

---

## Modeling transformed data

${log(\hat y_i)} = b_0 + b_1x_{i}$

Note that the linear model now has *log(y)* not just *y*.

We continue using `lm` as our main function to find the slope and y-intercept coefficients, but the argument is now `log(y) ~ x`.

`lm(log(y) ~ x, data = data_name)`

---

### Modeling log(mpg_city)

```{r}
model_mpg <- lm(log(mpg_city) ~ weight, data = cars)
model_mpg
```


${log(\hat y_i)} = b_0 + b_1x_{i}$

${log(\hat y_i)} = 4.192 + (-0.000355)x_{i}$

---

Use the model to estimated the city miles per gallon for a car that weighs 3,200 pounds.

${log(\hat y_i)} = 4.192 + (-0.000355)x_{i}$

${log(\hat y_i)} = 4.192 + (-0.000355)(3200)$

--

${log(\hat y_i)} = 3.056$ This result is for `log(y)`, so we have to solve for y to determine the mpg's.

--

$e^{log(\hat y_i)} = e^{3.056}$

--

$\hat y_i = e^{3.056}$ 

$\hat y_i = 21.24$

A car that weighs 3,200 pounds is expected to give you 21.24 city miles per gallon. 

---

### Interpretation of y-intercept and slope

${log(\hat y_i)} = b_0 + b_1x_{i}$

--

$e^{log(\hat y_i)} = e^{b_0 + b_1x_{i}}$

--

$\hat y_i = e^{b_0} \times e^ {b_1x_i}$


y-intercept: When $x = 0$, the expected response value (y) is $e^{b_0}$

Slope: For one-unit increase in x, the y is multiplied by $e^{b_1}$.

---

In our example, 

${log(\hat y_i)} = 4.192 + (-0.000355)x_{i}$

$\hat y_i = e^{4.192} \times e^ {-0.000355x_i}$

For one-pound increase in weight, the miles per gallon is multiplied by $e^{-0.000355}$.

The y-intercept cannot be interpreted as there is no such car that weighs 0 pounds.

---

### Other Common Transformations

  - $\sqrt{y}$
  - $\frac{1}{y}$
  - $y^n$
  
__If we transform our data using a function then we need to change the linear equation according to the function that we used.__ 