---
title: "Transformations"
author: "You name"
date: "Date here"
output: html_document

---


```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(broom)
library(openintro)

```

We know that in order to use a linear model there are certain conditions that needed to be met: 

  - linearity
  - constant variance
  - normality of residuals 

What if the condition of linearity is not met? 

We can transform the data so that the relationship gets close to a linear trend.

---

#### Data `car93` in `openintro` package

```{r}
cars <- cars93
glimpse(cars)
```

## Example

Let's say we wanted a model to predict city miles per gallon using the weight of the car. 

| y | Response    | weight  | Numeric |
|---|-------------|---------|---------|
| x | Explanatory | mpg_city| Numeric |


```{r message=FALSE, out.width = "50%"}
 
  
```

The heavier the car, the fewer miles per gallon you get, but they do not have a linear association. It's curved!

- Linear regression is not appropriate. 

### Using residual plot to verify linearity condition

- Run model:
```{r message = FALSE, warning = FALSE}

```

- Add residuals to data: 
```{r message = FALSE, warning = FALSE}

```

- Create residual plot:
```{r message = FALSE, warning = FALSE, out.width = "50%"}

```

The residual plot reveals a clear curve trend, hence a **linear model is not appropriate**. 

What can we do then?


---

### Transformations of response variable (*y*)

When the relationship between an explanatory variable *x* and the response variable *y* is not linear and is curved, then one thing we can do is transform the data of *y*. 

The transformed data then lessens the curve and approaches it closer to a linear relationship with the explanatory variable.  

One of the most common transformations is done by taking the log of the response variable values, `log(y)`. 

  - Note that log is natural log in R.
  - Data values cannot be negative. 
  - Check data for 0's and negative values. 
  - If there are a few 0's then we can add a small amount before taking the log, e. g. `log(y + 1)`.
  - If y has many zero values, or any negative values, a log-transformation might not be the best choice.
  


In our example, *y* = mpg_city, so we would take the log of `mpg_city` and plot it on the y-axis.


```{r }

```


### Modeling transformed data

${log(\hat y_i)} = b_0 + b_1x_{i}$

Note that the linear model now has *log(y)* not just *y*.

We continue using `lm` as our main function to find the slope and y-intercept coefficients, but the argument is now `log(y) ~ x`.

`lm(log(y) ~ x, data = data_name)`

---

### Modelin log(mpg_city)

```{r}

```


${log(\hat y_i)} = b_0 + b_1x_{i}$

${log(\hat y_i)} = ______ + (_______)x_{i}$


Use the model to estimate the city miles per gallon for a car that weighs 3,200 pounds.

${log(\hat y_i)} = 4.192 + (-0.000355)x_{i}$

${log(\hat y_i)} = 4.192 + (-0.000355)(3200)$

${log(\hat y_i)} = 3.056$ This result is for `log(y)`, so we have to solve for y to determine the mpg's.

$e^{log(\hat y_i)} = e^{3.056}$

$\hat y_i = e^{3.056}$

$\hat y_i = ________$

A car that weighs 3,200 pounds is expected to give you _______ city miles per gallon. 

---

### Interpretation of y-intercept and slope

${log(\hat y_i)} = b_0 + b_1x_{i}$

$\hat y_i = e^{b_0} \times e^ {b_1x_i}$


y-intercept: When ______, the expected response value (y) is _______

Slope: For one-unit increase in ______, the ______ is multiplied by _______

---

In our example,

${log(\hat y_i)} = 4.192 + (-0.000355)x_{i}$

$\hat y_i = e^{4.192} \times e^ {-0.000355x_i}$

For one ________ increase in _______, the _______ is multiplied by ________.

The y-intercept cannot be interpreted as there is no such car that weighs 0 pounds.

---

### Other Common Transformations

  - $\sqrt{y}$
  - $\frac{1}{y}$
  - $y^n$
  
__If we transform our data using a function then we need to change the linear equation according to the function that we used.__ 