---
title: "Multiple Linear Regression"
author: "Alma Castro, Cypress College"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
    anchor_sections: FALSE
---

class: title-slide


<br>
<br>
.pull-right[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(openintro)
library(broom)

```

## Simple Linear Regression

So far, we have only considered the linear regression model in one explanatory variable.

$y_i = \beta_0 +\beta_1x_{i}  + \epsilon_i$

However, oftentimes in real life, more complex linear models are needed. There are many problems in which a knowledge of more than one explanatory variable is necessary in order to obtain a better understanding and a better prediction of a particular response. 


---

## Multiple Linear Regression Model

In general, we have data on `n cases` and `k explanatory variables`.

$y_i = \beta_0 +\beta_1x_{1i}  + \beta_2x_{2i} + ...+ \beta_kx_{ki} + \epsilon_i$

$\beta_0$ is the intercept  
$\beta_1$ is the slope for first explanatory variable  
$\beta_2$ is the slope for second explanatory variable
...
$\beta_k$ is the slope for kth explanatory variable

$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point

---

### Finding a Multiple Regression Model

To find the estimates of the $\beta_i$ coefficients, we continue to use the linear model function. But now *y* is a function of the sum of the explanatory variables. 

`lm(y ~ x_1 + x_2 + ... + x_k, data = data_name)`


---

#### Data `babies` in `openintro` package
```{r echo = FALSE}
glimpse(babies)
```


```{r}
babies <- babies |> 
  mutate(smoke = as.logical(smoke))
```

---

Let's look at the relationship between `birth weight` as the response, and `gestation` and `smoke` as the explanatory variables.

<div align = "center">

| y | Response    | Birth weight | Numeric |
|---|-------------|-----------------|---------|
| $x_1$ | Explanatory | Gestation           | Numeric |
| $x_2$ | Explanatory | Smoke           | Categorical |


---

Scatter plot of `gestation (x_1)` and `birthweight (y)`, grouping points with color using `smoke (x_2)`.

.pull-left[
```{r eval = FALSE, warning = FALSE}
babies |> 
  ggplot(
    aes(x = gestation, 
        y = bwt, 
        color = smoke)) +
  geom_point()
```

]

.pull-right[
```{r echo = FALSE, warning = FALSE}
babies |> 
  ggplot(aes(x = gestation, y = bwt, color = smoke)) +
  geom_point()
```
]


---

## Regression Model for birthweight

$y_i = \beta_0 +\beta_1x_{1i}  + \beta_2x_{2i} + \epsilon_i$

$y_i$ is birthweight

$\beta_0$ is the intercept  
$\beta_1$ is the slope for **gestation**  
$\beta_2$ is the slope for **smoke** 

$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point

---

## Regression Model for birthweight

```{r}
model_gs <- lm(bwt ~ gestation + smoke, data = babies)
model_gs
```

--

$\hat {\text{bwt}_i} = b_0 + b_1 \text{ gestation}_i + b_2 \text{ smoke}_i$

$\hat {\text{bwt}_i} = -0.932 + 0.443\text{ gestation}_i + (-8.088)\text{ smoke}_i$

---

### Making Predictions

$\hat {\text{bwt}_i} = - 0.932 + 0.443\text{ gestation}_i + (-8.088)\text{ smoke}_i$

Find the expected birth weight for a baby who had 250 days of gestation with a **smoker** mother.

$\hat {\text{bwt}_i} = -0.932 + (0.443 \times 250) + (-8.088 \times 1) = 101.7$ ounces

Find the expected birth weight for a baby who had 250 days of gestation with a **non-smoker** mother.

$\hat {\text{bwt}_i} = -0.932 + (0.443 \times 250) + (-8.088 \times 0) = 109.8$ ounces


For all pregnancies with a gestation of 250 days, the difference in birth weight between babies of smokers and non smoker mothers is 8 ounces. 

---

Overall, the difference in birth weight between the two groups is about 8 ounces when controlling for gestation period. 
.pull-left[
```{r eval = FALSE, warning = FALSE}
babies |> 
  ggplot(aes(x = gestation, 
             y = bwt, 
             color = smoke)) +
  geom_point() +
  geom_abline(slope = 0.443, 
              intercept = -0.932, 
              color = "red") +
  geom_abline(slope = 0.443, 
              intercept = -9.02, 
              color = "blue")
```
]

.pull-right[

```{r echo = FALSE,warning = FALSE}
babies |> 
  ggplot(aes(x = gestation, y = bwt, color = smoke)) +
  geom_point() +
  geom_abline(slope = 0.443, intercept = -0.932, color = "red") +
  geom_abline(slope = 0.443, intercept = -9.02, color = "blue")
```

]
---

### Making predictions for all existing values in the data set 

```{r}
babies |> 
  modelr::add_predictions(model_gs) |> 
  select(bwt, gestation, smoke, pred)
```

---

And we can also add the residuals. 

```{r}
babies |> 
  modelr::add_predictions(model_gs) |> 
  modelr::add_residuals(model_gs) |> 
  select(bwt, gestation, smoke, pred, resid)
```


```{r eval = FALSE}
babies<- babies |> 
  mutate(smoke = case_when(
    smoke == 0 ~ FALSE,
    smoke == 1 ~ TRUE),
    smoke = as.factor(smoke))
glimpse(babies)
```



```{r eval = FALSE, message = FALSE, warning = FALSE}
model_gs |> 
  ggplot(aes(x = gestation, y = bwt, color = smoke)) +
  geom_point() + 
geom_smooth(method = "lm", se = FALSE)
```


---
## Inference: Confidence Interval

```{r}
confint(model_gs, level = 0.95)
```

--

Neither of the confidence intervals for the slopes contain 0.

We can conclude that **gestation has a positive relationship with birth weight and smoke has a negative relationship with birth weight.**

