---
title: "Multiple Linear Regression"
author: "Your Name"
date: "date here"
output: html_document
---


```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(openintro)
library(broom)
library(janitor)
```

## Simple Linear Regression 
So far, we have only considered the linear regression model in one explanatory variable.

$y_i = \beta_0 +\beta_1x_{i}  + \epsilon_i$

However, oftentimes in real life, more complex linear models are needed. There are many problems in which a knowledge of more than one explanatory variable is necessary in order to obtain a better understanding and a better prediction of a particular response. 

---

## Multiple Linear Regression Model

In general, we have data on `n cases` and `k explanatory variables`.

$y_i = \beta_0 +\beta_1x_{1i}  + \beta_2x_{2i} + ...+ \beta_kx_{ki} + \epsilon_i$

$\beta_0$ is intercept  
$\beta_1$ is the slope for first explanatory variable  
$\beta_2$ is the slope for second explanatory variable

$\beta_p$ is the slope for kth explanatory variable

$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point

### Finding a Multiple Regression Model

To find the estimates of the $\beta_i$ coefficients, we continue to use the linear model function. But now *y* is a function of the sum of the explanatory variables.

`lm(y ~ x_1 + x_2 + ... + x_k, data = data_name)`

---

#### Data `babies` in `openintro` package

```{r echo = FALSE}
glimpse(babies)
```


Let's look at the relationship between `birth weight` as the response, and `gestation` and `smoke` as the explanatory variables.


| y | Response    | Birth weight | Numeric |
|---|-------------|-----------------|---------|
| $x_1$ | Explanatory | Gestation           | Numeric |
| $x_2$ | Explanatory | Smoke           | Categorical |


Scatter plot of `gestation (x_1)` and `birthweight (y)`, grouping points with color using `smoke (x_2)`.

```{r}

```


### Regression Model for birthweight

$y_i = \beta_0 +\beta_1x_{1i}  + \beta_2x_{2i} + \epsilon_i$

$y_i$ is birthweight

$\beta_0$ is the intercept  
$\beta_1$ is the slope for **gestation**  
$\beta_2$ is the slope for **smoke** 

$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point



```{r}

```


$\hat {\text{bwt}_i} = b_0 + b_1 \text{ gestation}_i + b_2 \text{ smoke}_i$

$\hat {\text{bwt}_i} = _________ + ________\text{gestation}_i + (_________)\text{smoke}_i$


### Making predictions


Find the expected birth weight for a baby who had 250 days of gestation with a **smoker** mother.

$\hat {\text{bwt}_i} = -0.932 + (0.443 \times 250) + (-8.088 \times 1) = ________$

Find the expected birth weight for a baby who had 250 days of gestation with a **non-smoker** mother.

$\hat {\text{bwt}_i} = -0.932 + (0.443 \times 250) + (-8.088 \times 0) = ________$


For all pregnancies with a gestation of 250 days, the difference in birth weight between babies of smokers and non smoker mothers is _______ ounces. 

Overall the difference in birth weight between the two groups is about 8 ounces when controlling for gestation period.

```{r}

```


### Making predictions for all existing values in the data set

```{r}

```


And we can also add the residuals.

```{r}

```



## Inference: Confidence Interval

```{r}

```

Neither of the confidence intervals for the slopes contain _____.

We can conclude that ____________________________

