---
title: "Summarize Data Avtivity KEY (10 pts)"
author: "name here"
date: "date here"
output: html_document
---


```{r message = FALSE}
library(tidyverse)
library(janitor)
```


Use the `mpg` data to answer the following questions. Use the ?mpg to find more information about the data set and its variables.

1. (2 pts) Find the mean and median for highway mpg. Use the `summarize()` function.
```{r}
summarize(mpg, 
            mean(hwy),
            median(hwy))
```


2. (2 pts) Construct a relative frequency distribution for the type of car. What percent of the cars in the data set are SUVs?
```{r}
tabyl(mpg, class)

```

3. (2 pts) Find P_90 for the highway mpg variable. Interpret result in the context of the problem.
```{r}
quantile(mpg$hwy, c(0.9))
```


4. (2 pts) Find the first and third quartiles for the highway mpg variable. Interpret the third quartile in the context of the problem.
    
```{r warning = FALSE}
mpg %>% 
  summarize(quantile(hwy, c(0.25, 0.75)))
```

5. (2 pts) Find the standard deviation mpg in highway, both using the formula and with R. 
  a. step-by-step with formula
```{r}
xbar <- mean(mpg$hwy)
xbar
sqr_dist <- (mpg$hwy - xbar)^2
sum_dev <- sum(sqr_dist)
sd_hwy <- sqrt(sum_dev/(nrow(mpg)-1))
sd_hwy
```

  b. using R.
```{r}
sd(mpg$hwy)
```

