---
title: "Summarizing Data Numerically"
author: "Alma Castro, Cypress College"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css"]
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

```{r child = "setup.Rmd"}
```


```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(janitor)
options(scipen = 999)
```


<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`

]


---

class: inverse middle

```{r echo = FALSE, message = FALSE}
lapd <- 
  read_csv(here::here("Police_Payroll.csv")) |> 
  clean_names() |>
  filter(year == 2018) |> 
  select(job_class_title, 
         employment_type, 
         base_pay) |>
  mutate(employment_type = as.factor(employment_type),
           job_class_title = as.factor(job_class_title))
```



.font75[Summarizing Categorical Data]


---

class: middle

Categorical data are summarized with **counts** or **proportions**.

In statistics, summarized categorical data based on counts (frequencies) is called a __frequency distribution__. If it is based on proportions (relative frequencies) then it is called a __relative frequency distribution__.

---

## Frequency Distribution

To create a quick frequency distribution with the counts(frequencies) for each category in the categorical variable, we use the function `count(dataset, variable)` from the dplyr package.  

```{r}
count(lapd, employment_type)
```

---

## Relative Frequency Distribution

A relative frequency distribution displays the proportion of observations for each category of a categorical variable. 

$$relative frequency = \frac{frequency}{total frequencies}$$
---

To create a more detailed summary that include both the count and relative frequencies, we can use the `tabyl(dataset, variable)` function from the janitor package.


```{r}
tabyl(lapd, employment_type)
```

---

class: inverse middle

.font75[Summarizing Numerical Data]

---

class: middle

## Measures of Center

**Mean**: The average of a list of data values. The mean for a sample of size `n` is denoted with "x bar".

$$\bar x = \frac{\Sigma_{i = 1}^{n} x_i}{n}$$

**Median**: The middle value of an ordered data set.

**Mode**: The value that appears the most often (local maxima for continuous variables). 

  - A set of data can have no mode, one mode, or more than one mode. You can find the mode for quantitative and qualitative data.

---

class: middle

To summarize numerical data we can use the function `summarize()`, with the dataset as the first argument and the function of the numerical summary we want to compute as subsequent arguments.

`summarize(data, function(variable))`
  - where the `function` can be a summary function like `mean()`, `median()`...
  - The output will be a tibble. 
  
We can also simply use the function of the numerical summary we need with the input `data$variable` as the argument.

`function(data$variable)`
  - The output will be a single value.

---

## Mean

.pull-left[
```{r}
summarize(lapd, mean(base_pay))
```
]

--

.pull-right[

```{r}
mean(lapd$base_pay)
```

]

---

## Median

.pull-left[
```{r}
summarize(lapd, median(base_pay))
```
]

--

.pull-right[

```{r}
median(lapd$base_pay)
```

]

---

We can calculate multiple statistics and use multiple functions inside the summarize function, and we can add names to the output.

```{r}
summarize(lapd,
          minimum_base_pay = min(base_pay),
          mean_base_pay = mean(base_pay),
          med_base_pay = median(base_pay),
          maximum_base_pay = max(base_pay))
```

---

## Mode

```{r}
count(lapd, base_pay, sort = TRUE)
```

We don't use the function mode since it will only return the type of variable. 


---

## Measures of Position

**Quantiles**: cut points (values) that divide an ordered data set into equal parts, having equal probabilities. To find them, we specify the quantiles we want inside the function `quantile()` in the `summarize()` function. Commonly used quantiles: percentiles and quartiles. 

- __Percentiles__ divide the data into 100 equal parts. The kth percentile, denoted P_k, of a set of data is a value such that k percent of the observations are less than or equal to the value.

- __Quartiles__ divide the data into 4 equal parts (quarters). The quartiles are the 25th, 50th, and 75th  percentiles, where Q1 = 25th percentile, Q2 = 50th percentile, i.e. the median, and Q3 = 75th percentile. 

---

## Quartiles


- First quartile (Q1): 25% of the data falls at or below this value, and 75% falls above it.

- Median (Q2): 50% of the data falls at or below this value, and 50% falls above it - this is the middle value we discussed earlier.

- Third quartile (Q3): 75% of the data falls at or below this value, and 25% falls above it.


---

## Quartiles


```{r warning = FALSE}
summarize(lapd, quantile(base_pay, c(0.25, 0.50, 0.75)))
```

```{r echo = FALSE}
q1 <- summarize(lapd, quantile(base_pay, c(0.25))) %>%
  round(0) %>% 
  pull()
```


We would expect 25% of the data to be less than or equal to `r q1`.

Context: We expect 25% of employees at the Los Angeles Police Department to have a base pay less than or equal to $67,266.
