---
title: "Intro to Data Activity"
author: "Your name here"
date: "date"
output: html_document
---

```{r message=FALSE}
library(dplyr)
library(fivethirtyeight)
library(tidyverse)
data(candy_rankings)

```


Answer the following questions about the `candy_rankings` data frame using functions for data frames when appropriate. Submit html file to corresponding Canvas assignment. 

1. Use the help feature to find more information about the variables in the data set. Display the first few rows of the data set. 
```{r}
?candy_rankings
head(candy_rankings)
```


2. What does the variable "pluribus" describe? 

Describes whether a candy comes in a bag or a box. 


3. How many observations are there in this data set?
```{r}
nrow(candy_rankings)
glimpse(candy_rankings)
```


4. How many variables are there in this data set? Name 3 of them.
```{r}
ncol(candy_rankings)
glimpse(candy_rankings)

```

Three of the variables are: chololate, fruity, and caramel.

5. How many rows and columns does the data set have?

There are 85 rows and 13 columns.

