---
title: "Hypothesis Test Activity (20 pts) "
author: "name here"
date: "date here"
output: html_document

---

```{r echo = FALSE, message=FALSE, warning = FALSE}
library(tidyverse)
library(openintro)
library(titanic)
library(janitor)
data(mpg)
options(scipen=999) # to avoid scientific notation
titanic <- titanic_train |> 
  clean_names() |> 
  select(survived, pclass, sex, fare) |> 
  mutate(pclass = case_when(pclass == 1 ~ "First", 
                            pclass == 2 ~ "Second",
                            pclass == 3 ~ "Third"),
         sex = as.factor(sex))

```

1. Using the `mpg` data as a sample, is there enough evidence to determine that the average mpg in highway of compact cars in the past is less than the average 30 mpg of compact cars nowadays? Use a significance level of 1%.

a) (2 pts) Filter the data to contain information for compact cars only and save it under "comp". 
```{r}

```


b) (2 pts) State the null and alternative hypotheses.


c) (2 pts) Check conditions. 


d) (2 pts) Ignoring outliers, find the test statistic and the p-value. Round to 4 decimal places.
```{r}

```


e) (2 pts) Make a decision to reject the null hypothesis and write conclusion in the context of the problem.




2. Using the `titanic` data, and assuming conditions are satisfied, is there enough evidence to conclude that surviving and class are independent? Test at the 5% level of significance.

a) (2 pts) Find the contingency table.

b) (2 pts) What are the expected counts?

c) (2 pts) State the hypotheses.

d) (2 pts) Find the test statistic and p-value.

e) (2 pts) Make a decision and write conclusion in the context of the problem.


