---
title: "Hypothesis Testing: One Mean"
author: "Name here"
date: "date here"
output: html_document
---


```{r echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)

```


To **test hypotheses** regarding a population **mean**, assuming the population standard deviation **σ is unknown**, we use the t-distribution.

## Student's t Distribution 
Suppose an independent sample of size *n* is taken form a population that is normally distributed. Then the statistic 
$$t = \frac{\bar{x}-\mu}{\frac{s}{\sqrt{n}}}$$
follows a student's t distribution with *n-1* degrees of freedom. 

The t-distribution is also bell shaped, but its tails are thicker than the normal distribution’s, and its bell curve fluctuates depending on the sample size. 

---

## Steps of Hypothesis Testing

**Step 1**: Determine **variable** and **parameter** of interest in the context of the problem.

  - x = random variable
  - $\mu$ = Mean of random variable 


**Step 2**: State **hypotheses** regarding the mean based on the claims and determine type of test.

Identify $\mu_0$ to be the assumed value of the mean based on the claim being made from the context of the problem. $H_a$ is also determined by the claim being made.

#### Two-tailed test

  - $H_0 : \mu = \mu_0$  
  - $H_a : \mu \neq \mu_0$
  - Key words: is not, is different, has changed

  
#### Left-tailed test

  - $H_0 : \mu = \mu_0$  
  - $H_a : \mu < \mu_0$
  - Key words: less than, decreased, smaller
  
#### Right-tailed test

  - $H_0 : \mu = \mu_0$  
  - $H_a : \mu > \mu_0$
  - Key words: more than, greater than, increased, exceeds, larger


**Step 3**: Verify requirements or **conditions**.

  - Independent sample or sample size is 5% or less of the population size, *n ≤ 0.05N*.
  - Sample has no outliers, and it is obtained from a population that is normally distributed OR the sample size *n ≥ 30*.


**Step 4**: Analyze collected data and use corresponding distribution to find the **test statistic and p-value.**

To check how different the observed data is from the expected/null mean, we compute a **test statistic**.

$$t_0 = \frac{\bar{x}-\mu_0}{\frac{s}{\sqrt{n}}}$$

We can find this value following the formula, but we will use a function to find it.

To find the test statistic and the p-value we use the function `t.test()` and corresponding code based on the type of test. Recall that the variable has to be a data vector, not a data frame.

#### Two-tailed test

  - $H_a : \mu \neq \mu_0$

t.test(variable, mu = $\mu_0$, alternative="two.sided")


#### Left-tailed test

  - $H_a : \mu < \mu_0$

t.test(variable, mu = $\mu_0$, alternative="less")

#### Right-tailed test

  - $H_a : \mu > \mu_0$

t.test(variable, mu = $\mu_0$, alternative="greater")

**Step 5**: **Compare** p-value with stated significance level $\alpha$ and decide to reject or fail to reject the null hypothesis. 

  - If P-value < $\alpha$, reject the null hypothesis, $H_0$.
  - If the P-value > $\alpha$ , fail to reject the null hypothesis, $H_0$.

**Step 6**: Write a **conclusion in the context** of the problem. 

  - $H_0$ is rejected: There is sufficient evidence to conclude “alternative hypothesis in context.” 
  - $H_0$ is not rejected: There is insufficient evidence to conclude “alternative hypothesis in context.”

---

## Example 1

Suppose a manufacturer of a battery claims the mean life of the battery is 50 days. The quality control manager at a tech company that relies on that battery thinks this claim is inflated and that the actual average is lower.  

The manufacturer stands by their claim and agrees for an independent agency to randomly select 30 batteries out of thousands made and look at their lasting life. the data is bellow. Is the data collected by the agency good evidence against the manufacturer's claim, allowing the tech company to end the contract? Use a significance level of 1%. 



```{r}
hours <- c(49.1, 49.5, 50.3, 49.2, 50.7, 49.0, 48.9, 49.5, 49.7, 48.7, 49.8, 48.7, 49.3, 50.4, 50.1, 51.2, 49.8, 49.2, 48.2, 50.2, 51.3, 49.4, 49.7, 50.0, 48.8, 48.9, 50.4, 49.0, 49.7, 48.6)

```


**Step 1**: Determine **variable** and **parameter** of interest in the context of the problem.

  - x = ______________
  - $\mu$ = __________________

--

**Step 2**: State **hypotheses** regarding the mean based on the claims and determine type of test.
  
### Left-tailed test

  - $H_0 : \mu ______$  
  - $H_a : \mu ______$


**Step 3**: Verify requirements or **conditions**.

  - sample size is 5% or less of the population size
  *30 ≤ 0.05(thousands)*.
  
  - Sample has no outliers, and the sample size *n ≥ 30*.

```{r }

```


**Step 4**: Analyze collected data and use corresponding distribution to find the **test statistic and p-value.**


```{r}

```

---

**test statistic:**
```{r}


```


**p-value** = ____________


**Step 5**: **Compare** p-value with stated significance level $\alpha$ and decide to reject or fail to reject the null hypothesis. 

  - P-value = __________ $\alpha=0.01$, hence _________.



**Step 6**: Write a **conclusion in the context** of the problem. 

  - $H_0$ is ___________: There is __________ evidence to conclude that ___________________.


Now that we know that the batteries last less than 50 hours, should the tech company cancel the contract? 

  - Statistically, there is evidence that the batteries do not last as long as the manufacturer says they should. 
  - However, based on this sample there are only half hour less on average that the batteries last. 
  - There may not be practical significance in this case. 30 minutes do not seem like a large difference. 
  - It seems that it might not be worth it to break the contract over 30 minutes. You might also look at how much it would cost to find a new manufacturer. These are also questions to consider before making any changes. 
  - What this discussion should show you is that just because a hypothesis has statistical significance does not mean it has practical significance. The hypothesis test is just one part of a research process.
