---
title: "Hypothesis Test: Categorical Data - Test for Independence"
author: "Alma Castro"
output: html_document

---

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(broom)
library(openintro)
theme_set(theme_gray(base_size = 18)) 

```


## The $\chi^2$ Test of Independence

We have tested a hypothesis about a mean, which relates to numerical data.  

What if a statement is about categorical data? More specifically, what if we wanted to know if two categorical variables were related in some way?
 
 - We can perform what is called a test for independence.

**The $\chi^2$ test of independence** is a statistical test used to determine whether there is a significant association between two categorical variables.

- It helps to assess whether the occurrence of one categorical variable is related to the occurrence of another categorical variable.

### Chi-squared Distribution

To test hypotheses regarding the relationship between two categorical variables we use the **Chi-squared distribution**.

If conditions are satisfied, the $\chi^2$ statistic 
$$
\chi^2 = \sum \frac{
(\text{observed count - expected count})^2
}{
\text{expected count}
}
$$
follows a Chi-square distribution with (r-1)(c-1) degrees fo freedom, where *r* is the number of rows, and *c* is the number of columns.


In general, the Chi-squared distribution is skewed to the right, but its shape varies depending on the degrees of freedom. It gets more symmetric as the degrees of freedom increase.


```{r echo = FALSE}
x_vals <- 0:20
df_vals <- seq(2,10,2)

df_dat <- map_df(df_vals, ~ tibble(
  l = paste(.),
  x = x_vals,
  y = dchisq(x_vals, .)
))

ggplot(df_dat, aes(x, y, color = factor(l, levels = df_vals))) +
  geom_line(data = df_dat, 
             aes(x, y, color = factor(l, levels = df_vals))) +
  labs(title = "Chi-square Distribution",
       subtitle = "various degrees of freedom",
       x = "",y = 'Density',color = "Degrees of Freedom")
```


---

## Contingency Tables

In order to determine an association between two categorical variables, we first need to have the data organized in a table.

A **contingency table** shows the joint distribution of two categorical variables. It displays the frequencies of the different combinations of categories for the two variables. 

- Variables are identified as the row variable and the column variable.
- Each cell gives the number of individuals that fall under two categories, one from each variable.
- Contingency tables are also called two-way tables.


### Creating a Contingency Table in RStudio

We can create a two-way contingency table in two ways depending on the way the data is given to us: 

1. Data is given as **raw data** where each row represents an observation.

  The function `table(data$row_variable, data$column_variable)` will count how many observations fall under each category. 
 
2. Data is already given as **counts in a two-way table**. 

  We enter data in the form of column vectors, x, y, z, …,  as many columns as the table has. Then use the function `data.frame(x,y,z,…)`, which will organize the data in the same way it is organized in the two-way table.

---

## Hypotheses for a Test of Independence

**Step 1:** The null and alternative hypotheses for a $\chi^2$ test of independence are as follows:

  - $H_0$: The row and column variables are not associated; they are independent
  - $H_a$: The row and column variables are associated; they are dependent



### Expected Counts

The idea behind this test is to compare observed counts to the expected counts if the null hypothesis were true (if the variables are independent).

To calculate the expected count for the cell in the $i^{th}$ row and $j^{th}$ column, compute
$$
\text{Expected Count}_{\,row \, i,\, \;col \, j } = 
\frac{
(\text{row}\,\; i \; \text{total}) \times (\text{column}\,\; j \; \text{total}) 
}{
\text{table total}
}
$$

In terms of counts, if there was no relationship, the observed counts should be similar to the expected counts. 


**Step 2:** Verify Requirements or Conditions.

- Independent observations (within the sample)
- Large samples: At least 5 expected counts in each cell


**Step 3:** Finding the Chi-squared **test statistic and the p-value**.

To check how different the observed counts are from the expected counts, we compute a test statistic known as the chi-square test static. 

**The $\chi^2$ test statistic** is found by computing the ratio of how far the observed counts are from the expected counts.

$$
\chi^2 = \sum_{ij} \frac{
(\text{observed count - expected count})^2
}{
\text{expected count}
}
$$

We can find this value following the formula, but we will use a function i R to find it. 

Recall, the p-value is the probability of observing a sample statistic at least as extreme as the one that has been observed if the null hypothesis was true.

Hence it is the area under the Chi-squared distribution to the right hand side of the test statistic. 


**To find the test statistic and the p-value** we use the function `chisq.test(contingency_table, correct = FALSE)`



**Step 4**: **Compare** p-value with stated significance level $\alpha$ and decide to reject or fail to reject the null hypothesis. 

  - If P-value < $\alpha$, reject the null hypothesis, $H_0$.
  - If the P-value > $\alpha$ , fail to reject the null hypothesis, $H_0$.


**Step 5**: Write a **conclusion in the context** of the problem. 

  - $H_0$ is rejected: There is sufficient evidence to conclude “alternative hypothesis in context.” 
  - $H_0$ is not rejected: There is insufficient evidence to conclude “alternative hypothesis in context.”

---

## Example 1

[Are Emily and Greg More Employable than Lakisha and Jamal? A Field Experiment on Labor Market Discrimination](https://www.nber.org/papers/w9873). 

```{r message=FALSE, echo = FALSE}
data(resume) #from openintro package
resume <- resume |> 
  mutate(received_callback = as.logical(received_callback)) |> 
  select(received_callback, race)

```

```{r}
glimpse(resume)
```


Research question: Is getting a callback associated with the race of the (fictitious) job candidate? Use significance level of 5%.

Explanatory variable: _______,  Response variable: _________ 


**Step 1.** State hypotheses

  - $H_0$: Receiving a callback and race are _________
  - $H_a$: Receiving a callback and race are _________


```{r}

```


We can use the function `addmargins()` to add the totals (sums) of the rows and columns to our contingency table.   

 
```{r}


```


**Step 2.** Verify conditions

  - We will assume observations are independent
  - Expected counts in each cell are more than 5

If we want to see the expected table, we can also do that with the chisq.test() function.

```{r}

```



**Step 3.** Find the test statistic and p-value


```{r}


```


**Step 4.** Compare p-value with stated significance level and decide to reject or fail to reject the null hypothesis. 

Using a significance level of $\alpha = 0.05$, we can see that the p-value of _______ is _________ than 0.05. Hence we __________ the null hypothesis in favor of the alternative.



**Step 5.** Write a conclusion in the context of the problem.

We have enough evidence to conclude that receiving a callback and race are dependent.  

---

## Example 2

A professor at a community college conducted a study to assess the effectiveness of delivering an introductory statistics course via traditional lecture-based method, online delivery, and hybrid instruction methods. The grades received for a random sample of students in each of the courses were tallied. 


| Grade | Traditional | Online | Hybrid |
|-------|-------------|-----------------|
| A | 36 | 39 | 24 |
| B | 52 | 55 | 66 |
| C | 57 | 68 | 90 |
| D | 46 | 38 | 41 |
| F | 46 | 54 | 31 |

At the 0.05 level of significance, test the claim that grade earned is independent of the type of course.


**Step 1.** State hypotheses

  - $H_0$: Grade earned is ____________ of the type of course.
  - $H_a$: Grade earned _________ on the type of course.


### data
Data is already given as counts in a two-way table, so we crate the table using vectors and the function `data.frame()`. 
```{r}

```



  
**Step 2.** Verify conditions

  - We will assume observations are independent
  - Expected counts in each cell are more than 5

```{r}

```


**Step 3.** Find the test statistic and p-value

```{r}

```

$\chi^2$ = _________

p-value = _________


**Step 4.** Compare p-value with stated significance level 

Using a significance level of $\alpha = 0.05$, we can see that the p-value of ________ is _________ than 0.05. Hence we ________ the null hypothesis in favor of the alternative.


**Step 5.** Write a conclusion in the context of the problem.

We have enough evidence to reject the claim that grade earned is independent of the type of course, and conclude that _________________. 
  
  
  