---
title: "Importing Datasets"
author: "Alma Castro, Cypress College"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css"]
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide


```{r echo = FALSE, message = FALSE}
library(tidyverse)
options(scipen = 999)
```


<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]


---

class:inverse middle

.font75[Importing Datasets]

---

## Setting Working Directory

Set working directory first to the folder where you have the dataset saved. 

- You can see your current RStudio working directory by running the function `getwd()`.

Once you know the working directory, you can leave as is or set it to what you want it to be:

- Go to "Session", then "Set Working Directory", then "Choose Directory".

If this still gives you an error that it cannot find the file, then you will also have to set the working directory for the `R Markdown` file. 

- Click on the drop down arrow next to "knit" icon, then "Knit Directory", then "Current Working Directory".

---

## Importing datasers with `read.csv()` from base R

The base R function `read.csv("dataset_file_name")` will import a csv file. Hence the dataset has to be saved under a .csv format. 

If names of variables have spaces, the function will plut a "." in place of the space. However, this does not follow tidyverse style: lower case with words separated by an underscore `_`. Hence you would have to take further steps to clean the data so that it fits tidyverse style. 

If the name of the dataset is long or you would like to change the name then you can assign it to your desired name. 

`data <- read.csv("dataset_file_name")`


---

### Example

```{r}
lapd <- read.csv("Police_Payroll.csv")
glimpse(lapd)
```

---

## Importing datasets with `read_csv()` from tidyverse

A more common function when using tidyverse packages and style is `read_csv()`. However, this function does not clean the variable names. It simply adds `' '` to the names.

---

### Example

```{r message = FALSE}
lapd <- read_csv("Police_Payroll.csv")
glimpse(lapd)
```

---

We will learn how to clean variable names and data in future topics.

```{r message = FALSE}
library(janitor)
lapd <- clean_names(lapd)
glimpse(lapd)
```

---

### Using variables within the imported dataset
  
In order to work with variables from the dataset, we must tell R what dataset the variable is from by using `dataset_name$variable`.

Or we can use the code `attach(dataset_name)` to set the dataset as our working dataset and be able to use the variables without having to recall the dataset every time.  

---


### Example

.pull-left[
Option 1

```{r}
mean(lapd$projected_annual_salary)
```
]

.pull-right[
Option 2

```{r}
attach(lapd)
mean(projected_annual_salary)
```
]




