---
title: "Writing Functions"
author: "Your name here"
date: "date here"
output: html_document
---

```{r message = FALSE}
library("tidyverse")
library(titanic)
library(janitor)
titanic <- titanic_train %>% 
  clean_names()
```


# Functions
There are two types of functions in R: 

- Functions that are already written commands in R or from libraries that you install. 

- Functions that the users create. These functions will
remain in the command memory of the software unless you delete them or overwrite them. 

The command to create a function is `function()`.


## Writing functions
There are three key steps to creating a new function:

  - You need to pick a __name__ for the function. 

  - You list the inputs, or __arguments__, inside the function. If we had more than one argument, then the call would look like function(x, y, z).

  - You place the code you have developed in the __body__ of the function, a { block that immediately follows function(...).

__Template:__ 
```{r eval = FALSE}
name <- function(arguments) {
  body
}
```


There are main groups of functions: 

- Vector functions take one or more vectors as input and return a vector as output.
- Data frame functions take a data frame as input and return a data frame as output.
- Plot functions that take a data frame as input and return a plot as output.

---

## Vector Functions

### Simple example

Using mathematical operations

```{r}

```


- Name: `double()`
- Argument: `number`
- Body: `number*2`


```{r}

```


```{r}

```



### Having multiple arguments

```{r}

```


```{r}

```


```{r}

```



### Another example

```{r}


```


```{r}

```



### Using a vector as the argument

```{r}

```


```{r}

```

```{r}

```


```{r}

```


```{r}

```

---

## Data Frame Functions

### Filtering Function

Suppose we wanted to filter passengers on the titanic based on the their travel class and create a separate data frame for each where the only information we keep is their name, age, and survival status. 

We can start with the filtering code for first class

```{r}

```



To filter passenger traveling on second class, we can use the previous code and only change pclass to 2.

```{r}

```


Whenever you catch yourself copying and pasting code and only modifying a tiny part (e.g. pclass), write a function to automate the process. The tiny part that needs the modification can be the argument of the function.

```{r}


```



Passengers traveling in first class. 

```{r}

```


Passengers traveling in second class.

```{r}


```


Because the filtering function creates a data frame, we can save it if we would like.

```{r}

```


---

## Ploting Functions

### Histogram Function

Imagine you wanted to create a lot of histograms with only varying binwidths. 

Let `bin` stand for binwidth.

```{r}

```


binwidth = 30
```{r}

```



binwidth = 100
```{r}

```


Lets change the previous function so that we can create a histogram with various bindwiths for any numerical variable in any dataframe.

Let's use `df` to stand for data frame and `var` stands for variable, and `bin` stand for binwidth.  

```{r}

```

We can use ggplot2 to create a plotting function because `aes()` is a data-masking function.

To pass a column name or expression supplied as an argument to your function to another data-masking function, you must "embrace" the argument name with double curly braces ({{ }}).

Embracing a variable tells ggplot to use the value stored inside the argument, not the argument as the literal variable name. { var } will make the function look inside of var rather than looking for a variable called var.

variable = age 
binwidth = 10

```{r}

```

variable = age
binwidth 15

```{r}

```


---

## Creating a new function:

  - Pick a name for the function; short if possible and describing what the function does.

  - List the inputs, or arguments, to the function inside function. If we had more than one argument, then the code would look like function(x, y, z).

  - Place the code you have developed in body of the function, a { block that immediately follows function(...).


Note the overall process: you can make the function after you had figured out how to make it work with a simple input. It’s easier to start with working code and turn it into a function; it’s harder to create a function and then try to make it work.