---
title: "Subseting Data Activity (20 pts)"
author: "Your name here"
date: "date here"
output: html_document
---

```{r message = FALSE}
library("tidyverse")
library("janitor")
```

Use the `lapd` data set to answer the following questions. 

1. (3 pts) Import the Police_Payroll.csv data file using `read_csv()` function, clean names to fit tidy style, and assign it to `lapd`, all in one r chunk of code. Then glimpse the data.
```{r}

```

2. (2 pts) How many LAPD staff members had a "Police Officer I" as their job class title in the year 2016?
```{r}

```


3. (3 pts) How many LAPD staff members had overtime pay between $5,000 and $10,000 and were employed full time? The output should only be the number.  
```{r}

```


4. (3 pts) How many LAPD staff were Police Detective I, II, or III and had a base pay of 100K or higher?
```{r}

```

5. (3 pts) Display only the job class title for those who spend an average health cost of more than $19,000 and have full time employment.
```{r}

```


6. (6 pts) Subset the LAPD data to only contain the variables `base_pay`, `projected_annual_salary` and `overtime_pay` for those with a job title of Police Officer II from the year 2013. Assign the filtered data to lapd_officers and glimpse it.  
```{r}

```


