---
title: "Data Wrangling"
author: "Your name here"
date: "date here"
output: html_document
---

```{r message = FALSE}
library("tidyverse")
library("janitor")
```

## Pipe Operator |> 

### What is the average of 4, 8, 16 approximately?

**Solution 1: Functions within Functions**

```{r}

```

The problem with writing functions within functions is that things will get messy and more difficult to read and debug as we deal with more complex operations on data.

**Solution 2: Creating Objects**

```{r}

```

The problem with creating many objects is that we will end up with too many objects in `Environment`.


**Solution 3: The (forward) Pipe Operator |> **

The pipe operator allows for the output of a function to be the argument of another function.

```{r}

```


The output of the first function is the first argument of the second function.

# Subsetting Data

### Import Data Set

Import the Police_Payroll.csv data file using `read_csv()` function and assign it to `lapd`. Glimpse the data.
```{r message = FALSE}
lapd <- read_csv("Police_Payroll.csv")
glimpse(lapd)
```


## Cleaning variable names to fit Tidyverse style

Recall, `clean_names()` is a function from the Janitor package. It will convert names to tidyverse style: all lower case and with underscores in place of spaces.

```{r}

```

### Importing data and cleaning names using `pipe operator |>`

```{r eval = FALSE, message = FALSE}


```


## Subsetting variables/columns

`select()` is used to select the variables we want to keep from the data frame. The first argument is the dataset followed by the list of variables we want to select. 

This is useful when working with large datasets that contain many variables, but you only need a few for your specific analysis. The resulting data frame would have the **chosen columns** and keep **all the rows**.

We can either enter the data frame as the first argument.

```{r}

```


or we can pipe the data frame.
```{r}

```


`select` can also be used to drop certain variables if used with a negative sign.

```{r}

```


or we can use the exclamation mark (!) with the `c()` function to list the ones we want to drop.

```{r}

```


### Selection helpers

These functions are used as an argument in the `select()` function to narrow it down to the specified variables you want.

`starts_with()`  
```{r}

```

`ends_with()`  
```{r}

```

`contains()` 
```{r}

```


## Subsetting observations/rows

The `slice()` function allows us to examine (subset) rows by their position (**row number**) in the dataset.

The data below includes all the rows from third to seventh. Including third and seventh.

```{r}

```


The `filter()` function allows us to subset rows based on **conditions**, while keeping all the columns.

We use relational and logical operators to create these conditions.

### Relational Operators in R


| Operator | Description              |
|----------|--------------------------|
| <        | Less than                |
| >        | Greater than             |
| <=       | Less than or equal to    |
| >=       | Greater than or equal to |
| ==       | Equal to                 |
| !=       | Not equal to             |



### Logical Operators in R

| Operator | Description |
|----------|-------------|
| &        | and         |
| &#124;   | or          |



### Examples 

Q1. According to [datausa.io](https://datausa.io/profile/geo/los-angeles-ca#:~:text=In%202018%2C%20Los%20Angeles%2C%20CA,%2462%2C474%2C%20a%203.78%25%20increase.) Los Angeles had a median household income of \$62,474 in 2018. How many LAPD staff members had a base pay higher than $62,474 in year 2018 according to this data?

```{r}

```


- We can pipe the filtered data into the `nrow()` function to only obtain the number of observations that satisfy the stated condition.

```{r}

```

According to this data, ________ LAPD staff members had a base pay higher than $62474 in the year 2018.

Q2. How many observations are available between 2013 and 2015 including 2013 and 2015?

```{r}

```

There are 40227 observations are available between 2013 and 2015 including 2013 and 2015.

Q3. How many LAPD staff were employed full time in 2018?

```{r}

```

There were _______ LAPD staff that were employed full time in 2018.

---

### Remarks when selecting variables and filtering observations

When selecting variables or filtering observations or both, there are two ways to go about saving the new filtered data: 

- assign it to a new data set name, keeping the original data set intact. 
- assign it to the same name, overriding the original data set.  

Also, when doing both, selecting variables and filtering observations, there is a correct order of doing it. We must **first filter** the observations and **then select** the desired variables. 

Suppose that we wanted to only focus on the year 2018, and the variables job_class_title, employment_type, and base_pay. Let's subset our data accordingly and assign it to lapd.

```{r}

```


Let's glimpse the new filtered data frame. 
```{r}

```

