---
title: "Aggregating Data"
author: "Your name here"
date: "date here"
output: html_document

---

```{r message = FALSE, warning = FALSE}

library("tidyverse")
library("janitor")
options(scipen = 999) #to change scientific notation to standard
```


Import `lapd` data set, clean names, subset data to only focus on the year 2018, and the variables `job_class_title`, `employment_type`, and `base_pay`, and change variables to their correct type. Then reassign it to `lapd`.

```{r message = FALSE}
lapd <- read_csv("Police_Payroll.csv") |> 
  clean_names() |> 
  filter(year == 2018) |> 
  select(job_class_title, 
         employment_type, 
         base_pay) |> 
  mutate(employment_type = as.factor(employment_type),
         job_class_title = as.factor(job_class_title))
```


Recall that categorical data can be summarized with counts or proportions. 


```{r}

```



Whereas, numerical variables can be summarized with means, medians, standard deviations, five-number summary, ...

```{r}

```


**But what if we wanted to find numerical summaries for each of the categories in a categorical variable?** 


---

## Aggregating Data by Groups

Often we want to calculate statistics for different groups in our data.

We can find numerical summaries for each category of a categorical variable using the function `group_by()` to first group observations based on their corresponding category and then finding the numerical summary we want using the `summarize()` function. 

`group_by(data, variable)`


### Example

What is the median salary for each employment type?

First, we group individuals based on their employment type.

```{r}

```


Then, we find the median salary for `base_pay`. 

```{r}

```


We can also remind ourselves how many staff members there were in each group using the function `n()`, but only works inside `summarize()` and `mutate()`.

```{r}

```


### Using a boxplot to compare the median salary per employment type  


```{r}

```





