---
title: "Intro to ggplot Activity KEY (16 pts)"
author: "Name"
date: "Date"
output: html_document
---

```{r message=FALSE }
library(tidyverse)
library(titanic)
library(janitor)
titanic <- titanic_train %>% 
  clean_names() %>% 
  select(survived, pclass, sex, age, fare, embarked) %>% 
  mutate(pclass = case_when(pclass == 1 ~ "First", 
                            pclass == 2 ~ "Second",
                            pclass == 3 ~ "Third"),
         embarked = case_when(embarked == "S" ~ "Southampton",
                              embarked == "C" ~ "Cherbourg",
                              embarked == "Q" ~ "Queenstown"),
         embarked = as.factor(embarked),
         sex = as.factor(sex),
         pclass = as.factor(pclass),
         survived = as.logical(survived))
```



Use the `titanic` data to answer the following questions. Use `?titanic_train` to find more information about the data set and its variables. 

1. (2 pts) Make a bar graph to determine the number of passengers that embarked at each of the destinations. What city had the most passengers that got on the Titanic?

```{r}
ggplot(data = titanic,
       aes(x = embarked)) +
  geom_bar()
```


2. (4 pts) Construct a histogram to look a the distribution for the age of the passengers. Pick a binwidth that you find appropriate and color it as you like. Describe the shape of the distribution. 

```{r warning = FALSE}
ggplot(data = titanic,
       aes(x = age)) +
  geom_histogram(binwidth = 8,
                 color = "red",
                 fill = "darkgreen")
```


3. (2 pts) Find the five-number summary for age. Describe the third quartile. 

```{r}

```


4. (1 pts) Find the IQR for age. 

```{r}

```


5. (3 pts) Find the lower and upper limits for age and determine if there are any outliers. 

```{r}

```


6. (2 pts) Construct a boxplot for age. Color it. 

```{r}
ggplot(titanic, 
       aes(y = age)) +
  geom_boxplot(color = "pink")
```



7. (2 pts) Pick a variable of interest and create a graph. 

```{r}
ggplot(data = titanic,
       aes(x = survived)) +
  geom_bar()
```

```{r}
ggplot(data = titanic,
       aes(x = sex)) +
  geom_bar()
```
 


