---
title: "Data Visualization: More ggplot"
author: "Alma Castro, Cypress College"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css"]
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide


<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]




```{r echo = FALSE, message=FALSE, warning = FALSE}
library(palmerpenguins)
library(tidyverse)
library(titanic)
library(janitor)
library(fivethirtyeight)
titanic <- titanic_train %>% 
  clean_names() %>% 
  select(survived, pclass, sex, age, fare, embarked) %>% 
  mutate(pclass = case_when(pclass == 1 ~ "First", 
                            pclass == 2 ~ "Second",
                            pclass == 3 ~ "Third"),
         embarked = case_when(embarked == "S" ~ "Southampton",
                              embarked == "C" ~ "Cherbourg",
                              embarked == "Q" ~ "Queenstown"),
         embarked = as.factor(embarked),
         sex = as.factor(sex),
         pclass = as.factor(pclass),
         survived = as.logical(survived))
```


---

## Data

```{r}
glimpse(penguins)
```

---

.left-panel[
### R
```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```

- gray background
- white gridlines
- small black points
- original variable names on axes
- no title
]

.right-panel[
### Default Scatter Plot

```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point() 
```

]

---

class: middle inverse

.font75[Changing size, shape, and color of *ALL* the points in a **scatter plot**]



---

We had talked about how we can use color, shape, and size to distinguish/group poins based on other variables.

But what if we simply wan to change the size, shape, and color of **ALL** the points in a plot to change the look?

We can do so by specifying: 

- `size = #` 
- `color = "color_you_want"`
- `shape = "shape_you_want"` 

inside the `geom_point()` function. 

---

.left-panel[

```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```


]

.right-panel[


```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 2) 
```

]

---

.left-panel[
### changing size
```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4) #<<
```

Default size of points is 2.
]

.right-panel[


```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4) 
```

]


---


.left-panel[
### changing color
```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, #<<
             color = "purple") #<<
```


]

.right-panel[


```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4,
             color = "purple") 
```

]

---

.left-panel[
### changing shape
```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, #<<
             color = "purple", #<<
             shape = "triangle") #<<
```
 
]

.right-panel[


```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") 
```

]

---

class: middle inverse

.font75[Labeling the axes and adding a title to a plot]

---


We can change the labels on the x and y axes and include a title on our plot by **adding** another layer using the function `labs()` then specifying the labels and title inside the function.

`labs(x = "Label of x-axis", y = "Label of y-axis",  title = "Title of Plot")` 

---

.left-panel[
### labels

```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)", #<<
       y = "Bill Length (mm)", #<<
       title = "Palmer Penguins") #<<
```
]

.right-panel[

```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins")
```
]

---

class: middle inverse

.font75[Changing the look of the rectangular coordinate system]

We can **add** another layer to our scatter plots to change the look of the coordinate's background. 

---

`theme_bw()`

.left-panel[
```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)", 
       y = "Bill Length (mm)", 
       title = "Palmer Penguins") +
  theme_bw() #<<
  
```

]

.right-panel[
```{r echo = FALSE, fig.height=5, warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins") +
  theme_bw()
```

]

---

`theme_gray()`

```{r echo = FALSE, fig.height=4, fig.align='center', warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins") +
  theme_gray()
```

- Note: this is the default theme.

---

`theme_classic()`

```{r echo = FALSE, fig.height=4, fig.align='center', warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins") +
  theme_classic()
```

---

`theme_dark()`

```{r echo = FALSE, fig.height=4, fig.align='center', warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins") +
  theme_dark()
```

---

`theme_minimal()`

```{r echo = FALSE, fig.height=4, fig.align='center', warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins") +
  theme_minimal()
```

---

`theme_void()`

```{r echo = FALSE, fig.height=4, fig.align='center', warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins") +
  theme_void()
```

- Note: labels on the axes are deleted even when included in the code.

---

class: middle inverse

.font75[Changing the font size and color of labels]

---

To change the __font size__ and **color** of the labels and the title we **add** another layer using the function `theme()` and indicate the desired text size # and color for of the title and axes. 

`theme(plot.title = element_text(size = #, color = "color"),`
        `axis.title = element_text(size = #, color = "color"),`
        `axis.text = element_text(size = #, color = "color")`

If we want to make all the labels the exact same font size and color then we we can do that by adding the code: 

`theme(text = element_text(size = #, color = "color_you _want"))`

---


```{r eval = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)", 
       y = "Bill Length (mm)", 
       title = "Palmer Penguins") +
  theme_bw() +
  theme(plot.title = element_text(size=30,  #<<
                                  color = "blue"), #<<
        axis.title = element_text(size = 20, #<<
                                  color = "purple"), #<<
        axis.text = element_text(size = 18, #<<
                                 color = "black")) #<<
  
```

---


```{r echo = FALSE, warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4, 
             color = "purple",
             shape = "triangle") +
  labs(x = "Bill Depth (mm)",
       y = "Bill Length (mm)",
       title = "Palmer Penguins") +
  theme_bw() +
  theme(plot.title = element_text(size=30, 
                                  color = "blue"),
        axis.title = element_text(size = 20,
                                  color = "darkgreen"),
        axis.text = element_text(size = 18,
                                 color = "black"))
```


---

Customization of color, labels, text size, and theme backgrounds also works for other graphs, not just scatter plots. 

.left-panel[

```{r eval = FALSE}
ggplot(data = titanic,
       aes(x = pclass)) +
  geom_bar(color = "green", #<<
           fill = "orange") + #<<
  labs(x = "Class Rankings", #<<
       y = "Passengers", #<<
       title = "Class Distribtion Among Titanic Passengers") + #<<
  theme_bw() + #<<
  theme(text = element_text(size= 18, color = "blue")) #<<
```
]

.right-panel[

```{r echo = FALSE, fig.height=5}
ggplot(data = titanic,
       aes(x = pclass)) +
  geom_bar(color = "green", 
           fill = "orange") + 
  labs(x = "Class Rankings", 
       y = "Passengers",
       title = "Class Distribtion Among Titanic Passengers") + 
  theme_bw() + 
  theme(text = element_text(size= 18, color = "blue")) #<<
```

]

---

```{r eval = FALSE}
?theme
```


---

class: middle inverse

.font75[Splitting the x-y grid based on a categorical variable]

- Splitting side-by-side barplots
- Splitting scatter plots 

An additional feature you can consider is facets, which is an added layer used to create separate panels (splits the grid) to distinguish groups.

---

How can we add a third categorical variable to a side-by-side barplot? 

.left-panel[


```{r eval = FALSE }
ggplot(data = penguins,
       aes(x = island, 
           fill = species)) + 
  geom_bar(position = "dodge")
```

]

.right-panel[

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5 }
ggplot(data = penguins,
       aes(x = island, 
           fill = species)) + 
  geom_bar(position = "dodge")
```

]

---

We can include an additional categorical variable to split plots based on the categories of the variable by **adding** a layer using the function `facet_grid()`.

- **Splitting the x-axis**: add `facet_grid(cols = vars(variable))`

.left-panel[

```{r eval = FALSE }
ggplot(data = penguins,
       aes(x = island, 
           fill = species)) + 
  geom_bar(position = "dodge") +
  facet_grid(cols = vars(sex)) #<<
```

This plot involves **3 categorical variables**, where a side-by-side bar plot is created for each category.

]

.right-panel[

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5 }
ggplot(data = penguins,
       aes(x = island, 
           fill = species)) + 
  geom_bar(position = "dodge") +
  facet_grid(cols = vars(sex))
```

]

---

We can also split scatter plots.

.left-panel[

```{r eval = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4,
             color = "purple")
```

]

.right-panel[
```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4,
             color = "purple")
```
]


---

This plot involves **2 numerical variables and 1 categorical variable**, where scatter plots are created for each category.

.left-panel[

```{r eval = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4,
             color = "purple") +
  facet_grid(cols = vars(sex)) #<<
```

]

.right-panel[
```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4,
             color = "purple") +
  facet_grid(cols = vars(sex))
```
]

---

This plot involves **2 numerical variables and 2 categorical variable**, where scatter plots are created for each category and points are grouped based on color. 

.left-panel[

```{r eval = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           color = island)) + #<<
  geom_point(size = 4) +
  facet_grid(cols = vars(sex)) #<<
```

]

.right-panel[
```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           color = island)) +
  geom_point(size = 4) +
  facet_grid(cols = vars(sex))
```
]

---


This plot involves **2 numerical variables and 3 categorical variable**, where scatter plots are created for each category and points are grouped based on color and shape. 

.left-panel[

```{r eval = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           shape = species, #<<
           color = island)) + #<<
  geom_point(size = 4) +
  facet_grid(cols = vars(sex)) #<<
```

]

.right-panel[
```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           shape = species,
           color = island)) +
  geom_point(size = 4) +
  facet_grid(cols = vars(sex))
```
]


---


This plot involves **3 numerical variables and 3 categorical variable**, where scatter plots are created for each category and points are grouped based on color and shape, and size of points are distinguished. 

.left-panel[

```{r eval = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           shape = species, #<<
           color = island, #<<
           size = body_mass_g)) + #<<
  geom_point() +
  facet_grid(cols = vars(sex)) #<<
```

]

.right-panel[
```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           shape = species,
           color = island,
           size = body_mass_g)) +
  geom_point() +
  facet_grid(cols = vars(sex))
```
]


---

class: middle 

We may split the grid horizontally instead of vertically. 
 
 - **Splitting the y-axis**: `facet_grid(rows = vars(variable))`

.left-panel[

```{r eval = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4,
             color = "purple") +
  facet_grid(rows = vars(sex)) #<<
```

]

.right-panel[
```{r echo = FALSE, fig.align='center', warning = FALSE, fig.height=5}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 4,
             color = "purple") +
  facet_grid(rows = vars(sex))
```
]

---

class: middle

You can do much more in ggplot. 

- Use images 
- Make maps
- Pick colors for every aspect of a graph
- Customize legends


We don't have time to cover all these. However, as we go along, you will become fluent in reading R code and Googling. 


Check out [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html#1) for some inspiration. 

For more on labels and chaging legends, see the Graphics for Cummunication section in the [R for Data Science](https://r4ds.had.co.nz/graphics-for-communication.html) book.

