---
title: "Data Visualization: More ggplot"
author: "Name"
date: "Date"
output: html_document
---

```{r echo=FALSE, message=FALSE }
library(palmerpenguins)
library(tidyverse)
library(titanic)
library(janitor)
titanic <- titanic_train %>% 
  clean_names() %>% 
  select(survived, pclass, sex, age, fare, embarked) %>% 
  mutate(pclass = case_when(pclass == 1 ~ "First", 
                            pclass == 2 ~ "Second",
                            pclass == 3 ~ "Third"),
         embarked = case_when(embarked == "S" ~ "Southampton",
                              embarked == "C" ~ "Cherbourg",
                              embarked == "Q" ~ "Queenstown"),
         embarked = as.factor(embarked),
         sex = as.factor(sex),
         pclass = as.factor(pclass),
         survived = as.logical(survived))
```


### Data

```{r}
glimpse(penguins)
```

---

### Default Scatter Plot

- gray background
- white gridlines
- small black points
- original variable names on axes
- no title


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point() 
```


## Changing size, shape, and color of *ALL* the points in a **scatter plot**

We had talked about how we can use color, shape, and size aesthetics to distinguish/group points based on other variables.

But what if we simply wan to change the size, shape, and color of **ALL** the points in a plot to change the look?

We may change the size, shape, and color of **ALL** the points in a plot by customizing them inside the `geom_point()` function.  

- `size = #` 
- `color = "color_you_want"`
- `shape = "shape_you_want"` 
 
### changing size of points

```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```


### changing color
```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```


### changing shape
```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```


---

## Labeling the axes of a scatter plot


We can change the labels on the x and y axes and include a title on our plot by **adding** another layer using the function `labs()` then specifying the labels and title inside the function.

`labs(x = "Label of x-axis", y = "Label of y-axis",  title = "Title of Plot")` 

```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```

---

## Changing the look of the rectangular coordinate system

We can **add** another layer to our scatter plots to change the look of the coordinate's background. 

- `theme_bw()`

```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
  
```



- `theme_gray()`
   - Note: this is the default theme.

```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
  
```


- `theme_classic()`

```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
  
```


- `theme_dark()`

```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
  
```


- `theme_minimal()`

```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
  
```


- `theme_void()`
    - Note: labels on the axes are deleted even when included in the code. 


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
  
```

---

## Changing the font size and color of labels

To change the __font size__ and **color** of the labels and the title we **add** another layer using the function `theme()` and indicate the desired text size # and color for of the title and axes. 

`theme(plot.title = element_text(size = #, color = "color"),`
        `axis.title = element_text(size = #, color = "color"),`
        `axis.text = element_text(size = #, color = "color")`

If we want to make all the labels the exact same font size and color then we we can do that by adding the code: 

`theme(text = element_text(size = #, color = "color_you _want"))`


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
  
```


Customization of color, labels, text size, and theme backgrounds also works for other graphs, not just scatter plots. 

```{r}
ggplot(data = titanic,
       aes(x = pclass)) +
  geom_bar(color = "green",
           fill = "orange") +
  labs(x = "Class Rankings",
       y = "Passengers",
       title = "Class Distribtion Among Titanic Passengers") +
  theme_bw() + 
  theme(text = element_text(size= 18, color = "blue"))
```



## Splitting the x-y grid based on a categorical variable

- Splitting side-by-side barplots
- Splitting scatter plots 

An additional feature you can consider is facets, which is an added layer used to create separate panels (splits the grid) to distinguish groups.

We can include an additional categorical variable to split plots based on the categories of the variable by **adding** a layer using the function `facet_grid()`.

- **Splitting the x-axis**: `facet_grid(cols = vars(variable))`

The plot below involves **3 categorical variables**, where side-by-side bar plots are created for each category.

```{r}
ggplot(data = penguins,
       aes(x = island, 
           fill = species)) + 
  geom_bar(position = "dodge")
```



### Splitting scatter plots.


- This plot involves **2 numerical variables and 1 categorical variable**, where scatter plots are created for each category.


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```



- This plot involves **2 numerical variables and 2 categorical variable**, where scatter plots are created for each category and points are grouped based on color. 


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```



- This plot involves **2 numerical variables and 3 categorical variable**, where scatter plots are created for each category and points are grouped based on color and shape. 


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```


- This plot involves **3 numerical variables and 3 categorical variable**, where scatter plots are created for each category and points are grouped based on color and shape, and size of points are distinguished. 


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```


We may **split the grid horizontally** instead of vertically. 
 
 - **Splitting the y-axis**: `facet_grid(rows = var(variable))`


```{r warning = FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) +
  geom_point()
```


---

You can do much more in ggplot. 

- Use images 
- Make maps
- Pick colors for every aspect of a graph
- Customize legends


We don't have time to cover all these. However, as we go along, you will become fluent in reading R code and Googling. 


Check out [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html#1) for some inspiration. 

For more on labels and chaging legends, see the Graphics for Cummunication section in the [R for Data Science](https://r4ds.had.co.nz/graphics-for-communication.html) book.

