---
title: "Homework 3: Study Design (22 pts)"
author: "Your name here"
date: "date here"
output: html_document
---

```{r message=FALSE}
library(openintro)
library(tidyverse)
resume <- resume |> 
  mutate(received_callback = as.logical(received_callback))
```

__The following questions 1 through 6 are related to the study [Are Emily and Greg More Employable than Lakisha and Jamal? A Field Experiment on Labor Market Discrimination](https://www.nber.org/papers/w9873). For now we will focus on the following part of the summary (abstract) of this study.__

>We perform a field experiment to measure racial discrimination in the labor market. We respond with fictitious resumes to help-wanted ads in Boston and Chicago newspapers. To manipulate perception of race, each resume is assigned either a very African American sounding name or a very White sounding name. The results show significant discrimination against African-American names: White names receive 50 percent more callbacks for interviews. 

In addition to the abstract authors mention the following in the paper:

>We randomly assign very White sounding names (such as Emily Walsh or Greg Baker) to half the resumes and very African American sounding names (such as Lakisha Washington or Jamal Jones) to the other half. Because we are also interested in how credentials affect discrimination, we experimentally vary the quality of the resumes used in response to a given ad. Higher quality applicants have on average a little more labor market experience and fewer holes in their employment history; they are also more likely to have an email address, have completed some certification degree, possess foreign language skills or have been awarded some honors. In practice, we typically send four resumes in response to each ad: two higher quality and two lower quality ones. We randomly assign to one of the higher and one of the lower quality resumes an African American sounding name.

The data from this study is provided as a data frame named `resume`. 
The variables that we are interested in are `received_callback` and `race` but you may also see in the data matrix that there are many other variables that the researchers controlled for (e.g. `years_experience`, `computer_skills` etc.)

Write your name and date, and answer the following questions using an r chunk when possible to show your code. 

## Question 1
(2 pts) How many observations (fictional job candidates) are in the resume data matrix? 
```{r}

```


## Question 2
(1 pts)This study is an

a. observational study
b. experimental study

## Question 3
(2 pts) What is the response variable?  


## Question 4
(2 pts) What is(are) the explanatory variable(s)?


## Question 5
The authors state:

>Applicants with White names need to send about 10 resumes to get one callback whereas applicants with African American names need to send around 15 resumes to get one callback. This 50 percent gap in callback rates is statistically very significant. Based on our estimates, a White name yields as many more callbacks as an additional eight years of experience. Since applicants’ names are randomly assigned, this gap can only be attributed to the name manipulation.

(2 pts) Do you think that authors are correct in stating that the significant difference in callback rates can only be attributed to the name manipulation (Black vs. White)?  Why or why not? 


## Question 6 
(3 pts) Take a random sample of 5 candidates. Use a seed of 220. 
```{r}

```


__The following questions 7 through 10 are based on this other scenario:__

>A faculty member in the English Department at Reagan High School wants to determine if the type of course taken, in-person vs online, has an effect on performance. Additionally, they think that there may be a difference in the performance of the men and women. The faculty member takes one of thier large class and, to accommodate any potential differences, they randomly assign half the men to the in-person class and the other half to the online version, and they do the same for the women. At the end of the semester, both classes will be given a final test to determine which group has a better mastery of the concepts in the course.

## Question 7
(4 pts) In this experiment, what is the response variable? What is the explanatory variable? 

## Question 8
(2 pts) What type of experiment is this? 

a. randomized experiment
b. randomized block experiment

Explain.


## Question 9
(2 pts) What are the treatments?

## Question 10
(2 pts) Can this experiment be done with blinding? Explain.
