---
title: "Logistic Regression Activity (24 pts)"
author: "name here"
date: "date here"
output: html_document
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(bayesrules)
library(tidyverse)
library(modelr)
library(broom)
data(weather_perth)

weather <- weather_perth %>% 
  select(day_of_year, raintomorrow, humidity9am, humidity3pm, raintoday)
```

```{r}
glimpse(weather)
```


Using the `weather` data from Perth, Australia, fit a logistic regression model where the response variable is `raintomorrow` and the explanatory variable is `humidity9am` (measured in hpa).

1. (2 pts) Visualize the data. 


2. (3 pts) Fit a logistic regression model, and write out the equation of your line. Round values to 3 decimals.

3. (2 pts) write model in odds form. 

4. (2 pts) write the model in probability form. 

5. (1 pts) What is the probability of rain tomorrow if humidity at 9 am is 0 hpa?  

6. (2 pts) Use result from question 5 to interpret the intercept coefficient.

7. (2 pts) What is the probability of rain tomorrow if humidity at 9 am is 1 hpa? 2 hpa? 

8. (2 pts) Use results from question 7 to interpret the slope coefficient. 

9. (2 pts) Add `raintoday`to the model from question 1. Write the equation of the model. Round values to 3 decimals. 

10. (2 pts) write model in probability form. 

11. (2 pts) What is the probability of rain tomorrow if humidity at 9 am is at 20 hpa and it did not rain today? 

12. (2 pts) Compare models and determining which one is best. 

