---
title: "Confidence Interval: Bootstrapping Activity KEY (12 pts)"
author: "name here"
date: "date here"
output: html_document

---

```{r echo = FALSE, message=FALSE, warning = FALSE}
library(tidyverse)
library(janitor)
library(infer) # for bootstrap related functions
options(scipen=999) # to avoid scientific notation

```


```{r message=FALSE, warning = FALSE}
lapd <- read_csv("Police_Payroll.csv") |> 
  clean_names() |> 
  filter(year == 2018) |> 
  select(base_pay)
```


1. (2 pts) Take a random sample of size 30 from base pay. Use seed of 2026.
```{r}
set.seed(2026)
lapd_samp <- sample_n(lapd, 30)
```


2. (4 pts) Take 1000 bootstrap samples of size 30, and find the median base pay. Use seed 2026.
```{r}
boot_median <- lapd_samp |>
  specify(response = base_pay) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "median")
boot_median
```


3. (2 pts)Visualize the bootstrap distribution. What shape does it have? 
```{r}
visualize(boot_median) +
  theme_bw()
```


4. (2 pts) Using the bootstrap samples from question 1, build a bootstrap 90% confidence interval to estimate the median base pay. Interpret your result. 
```{r}
boot_median |> 
  summarize(lower_bound = quantile(stat, 0.05),
            upper_bound = quantile(stat, 0.95))
```


5. (2 pts) What is the true median for the population. Does your interval contain the true median base pay? 
```{r}
median(lapd$base_pay)
```



