---
title: "Confidence Interval: Bootstrapping Activity (12 pts)"
author: "name here"
date: "date here"
output: html_document

---

```{r echo = FALSE, message=FALSE, warning = FALSE}
library(tidyverse)
library(janitor)
library(infer) # for bootstrap related functions
options(scipen=999) # to avoid scientific notation

```


```{r message=FALSE, warning = FALSE}
lapd <- read_csv("Police_Payroll.csv") |> 
  clean_names() |> 
  filter(year == 2018) |> 
  select(base_pay)
```



1. (2 pts) Take a random sample of size 30 from base pay. Use seed of 2026.
```{r}

```



2. (4 pts) Take 1000 bootstrap samples of size 30, and find the median base pay.
```{r}

```


3. (2 pts) Visualize the bootstrap distribution. What shape does it have? 
```{r}

```


4. (2 pts) Using the bootstrap samples from question 1, build a bootstrap 90% confidence interval to estimate the median base pay. Interpret your result. 
```{r}

```


5. (2 pts) What is the true median for the population. Does your interval contain the true median base pay? 
```{r}

```



