---
title: "Writing Functions"
author: "Alma Castro, Cypress College"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(titanic)
library(janitor)
titanic <- titanic_train |> 
  clean_names()
```


<br>
<br>
.pull-right[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

## Functions
There are two types of functions in R: 

- Functions that are already written commands in R or from libraries that you install. 

- Functions that the users create. These functions will
remain in the command memory of the software unless you delete them or overwrite them. 

The command to create a function is `function()`.

---

## Writing functions
There are three key steps to creating a new function:

  - You need to pick a __name__ for the function. 

  - You list the inputs, or __arguments__, inside the function. If we have more than one argument, then the call would look like `function(x, y, z)`.

  - You place the code you have developed in the __body__ of the function, a { } block that immediately follows function(...).

Template: 
```{r eval = FALSE}
name <- function(arguments) {
  body
}
```


---

There are main groups of functions: 

- Vector functions take one or more vectors as input and return a vector as output.
- Data frame functions take a data frame as input and return a data frame as output.
- Plot functions that take a data frame as input and return a plot as output.

---

class: inverse middle

.font75[Vector Functions]

---

### Simple example

Using mathematical operations

```{r}
double <- function(number) {
  number*2
}
```

--

- Name: `double()`
- Argument: `number`
- Body: `number*2`

---

```{r}
double(number = 3)
```

--

```{r}
double(3)
```


---

### Having multiple arguments

```{r}
multiply <- function(number, by) {
  number*by
}
```

--

```{r}
multiply(number = 10, by = 3)
```

--

```{r}
multiply(10, 3)
```


---
### Another example

```{r}
lm <- function(num) {
  2*num + 1
}
```

--

```{r}
lm(0)
lm(1)
```

---

### Using a vector as the argument

```{r}
a <- c(0, 1, 2, 3, 4, 5)
```

--

```{r}
double(a)
```

```{r}
lm(a)
```

---

```{r}
line <-data.frame(a, lm(a))
line
```

---

.pull-left[
```{r eval = FALSE}
ggplot(line,
       aes(x = a,
           y = lm(a)))+
  geom_point()
```
]

.pull-right[
```{r echo = FALSE}
ggplot(line,
       aes(x = a,
           y = lm(a)))+
  geom_point()
```
]

---

class: inverse middle

.font75[Data Frame Functions]

---

### Filtering Function

Suppose we wanted to filter passengers on the titanic based on the their travel class and create a separate data frame for each where the only information we keep is their name, age, and survival status. 

We can start with the filtering code for first class

```{r}
filter(titanic, pclass == 1) |>
    select(name, age, survived)
```

---

To filter passenger traveling on second class, we can use the previous code and only change pclass to 2.

```{r}
filter(titanic, pclass == 2) |>
    select(name, age, survived)
```

---

Whenever you catch yourself copying and pasting code and only modifying a tiny part (e.g. pclass), write a function to automate the process. The tiny part that needs the modification can be the argument of the function.

```{r}
passenger <- function(class) {
    filter(titanic, pclass == class) |>
    select(name, age, survived)
}
```

---

Passengers traveling in first class. 

```{r}
passenger(1)
```


---

Passengers traveling in second class.

```{r}
passenger(2)

```

---

Because the filtering function creates a data frame, we can save it if we would like.

```{r}
class3 <- passenger(3)
class3
```


---

class: inverse middle

.font75[Plotting Functions]
---

## Histogram Function

Imagine you wanted to create a lot of histograms with only varying binwidths. 

Let `bin` stand for binwidth.

```{r}
hist <- function(bin) {
    ggplot(titanic, 
           aes(x = fare)) + 
    geom_histogram(binwidth = bin)
}
```

---

.pull-left[
```{r}
hist(30)
```
]

--

.pull-right[
```{r}
hist(100)
```
]

---

Lets change the previous function so that we can create a histogram with various bindwiths for any numerical variable in any dataframe.

Let's use `df` to stand for data frame and `var` stands for variable, and `bin` stand for binwidth.  

```{r}
histogram <- function(df, var, bin) {
    ggplot(df, aes({{ var }})) + 
    geom_histogram(binwidth = bin)
}
```

We can use ggplot2 to create a plotting function because `aes()` is a data-masking function.

To pass a column name or expression supplied as an argument to your function to another data-masking function, you must "embrace" the argument name with double curly braces ({{ }} ).

Embracing a variable tells ggplot to use the value stored inside the argument, not the argument as the literal variable name. { var } will make the function look inside of var rather than looking for a variable called var.

---

```{r warning = FALSE}
histogram(df = titanic, var = age , bin = 10)
```

---

```{r warning = FALSE}
histogram(titanic, age, 15)
```


---

## Creating a new function:

  - Pick a name for the function; short if possible and describing what the function does.

  - List the inputs, or arguments, to the function inside function. If we had more than one argument, then the code would look like function(x, y, z).

  - Place the code you have developed in body of the function, a { } block that immediately follows function(...).


Note the overall process: you can make the function after you had figured out how to make it work with a simple input. It’s easier to start with working code and turn it into a function; it’s harder to create a function and then try to make it work.
